---
title: "Assignment 2"
output: html_document
---
###Scott Ogden

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Raw Data:
```{r}
raw.data <- raw.data<-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"  
```


###Problem 3:

Copy the introductory example. The vector "name"" stores the extracted names.

Name
```{r}
library("stringr")
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
```


a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name

```{r}
#Note: Used documentation and examples from http://www.endmemo.com/program/R/sub.php as a guide for replacement

#Remove titles
name_notitle<-sub("[A-z]{2,3}\\. ","",name)

#Swap names seperated by commas
name_swap <- sub("(\\w+),\\s(\\w+)","\\2 \\1", name_notitle)
#Now swap names seperated by period
name_swap2<-sub("(\\w+)\\s(\\w+).\\s(\\w+)","\\1. \\3 \\2", name_swap)

name_swap2

```

b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.)

```{r}
#Use the remove title logic and "Detect String" Function
title<-str_detect(name,"[A-z]{2,3}\\. ")
#Join
data.frame(name_swap2,title)
```

c) Construct a logical vector indicating whether a character has a second name

```{r}
count <- str_count( name_swap2, "\\S+" )
data.frame(name_swap2,count>2)

```

###Problem 4
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

(a) [0-9]+\\$

This will find numeric strings that have a dollar sign attached to the end.
```{r}
a <- c("1234asd","12345$","$123as","sadsf$")
str_detect(a, "[0-9]+\\$")
```



(b) \\b[a-z]{1,4}\\b

This will find the strings that have from exactly four lower case letters in them
```{r}
b <- c("abcde","123aabc22$","Abcd","abcd")
str_detect(b, "\\b[a-z]{1,4}\\b")
```



(c) .*?\\.txt$

This will find strings that end in .txt. Because of the ? the preceeding txt is optional so it will find anything ending .txt.
```{r}
c <- c("1234asd.txt",".txt","sa.txtdk$$","$")
str_detect(c, ".*?\\.txt$")
```



(d) \\d{2}/\\d{2}/\\d{4}

This will find strings that are formatted like numbers seperated by "/".  Specifically, strings that have at least two digits followed by / then exactly two digits then / then exactly 4 digits.

```{r}
d <- c("aa/bb/cccc","09/18/2016","9/18/2016","9999999/99/9999")
str_detect(d, "\\d{2}/\\d{2}/\\d{4}")
```



(e) <(.+?)>.+?</\\1>

This will find strings that are enclosed by the following: <ANYTHING> ANYTHING ELSE </ANYTHING>.  Basically carrots containing a some text contain some other string, much like html tags, with the appropriate "/" in the closing tag.
```{r}
e <- c("<b> Hi </b>","asdasfa","<i>Italic</i>","<i>Failed<i>")
str_detect(e, "<(.+?)>.+?</\\1>")
```


###Problem9

The following code hides a secret message. Crack it with R and regular expressions.

```{r}
code<-"clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

# First find all capital letters
caps<-str_extract_all(code, "[[:upper:].]{1,}")
#unlist
u<-unlist(caps)
#concatenate all
pst<-paste(u, collapse ='')
#replace periods
result<-str_replace_all(pst,"[.]"," ")
result

```

