---
title: "Week 7 Assignment"
author: "Scott Ogden"
date: "October 16, 2016"
output: html_document
---

```{r setup, include=TRUE}
library("XML")
library(plyr)
library(knitr)
```

### Assignment

In this assignment we will load data from three different sources into R (as a prelude to further analysis)

### Part 1: html

Data from:
https://github.com/scottogden10/607-Assignment2/blob/master/Books.html

Use the htmltab library to load in the values and format as dataframe.

```{r}
html<-"https://raw.githubusercontent.com/scottogden10/607-Assignment2/master/Books.html"
library(htmltab)

htmldf<-htmltab::htmltab(html)
htmldf

```


### Part 2: XML

Data from:
https://github.com/scottogden10/607-Assignment2/blob/master/Books.xml

Use the XML and Rcurl libraries to load in the values and format as dataframe.

```{r}
xml<-"https://raw.githubusercontent.com/scottogden10/607-Assignment2/master/Books.xml"

library("XML")
library(RCurl)
xmldata<-getURL(xml)
xmlparse<-xmlParse(xmldata)
xmlparse

xmldf<-xmlToDataFrame(xmldata)
xmldf
```

### Part 3: json

Data from:
https://github.com/scottogden10/607-Assignment2/blob/master/Books.json

Use the jsonlite library to load in the values and format as dataframe.

```{r}
library(jsonlite)

json<-"https://raw.githubusercontent.com/scottogden10/607-Assignment2/master/Books.json"
jsonurl=getURL(json)
jsonlist<-jsonlite::fromJSON(jsonurl,simplifyDataFrame = TRUE)
jsondf<-data.frame(jsonlist)

jsondf
```

Note all the dataframes are identical with this code!
